---
title: "Marginal Effects"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    css: style.css
    theme: paper
    source_code: "https://github.com/josherrickson/marginsnotes"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Sidebar {.sidebar}
-------------------------------------

### Layout

On the left sidebar for each page, you'll find a (mostly) software-agnostic discussion of the topic at hand.

The main panels of each page consist of a tabbed series of pages with the code and output in the noted language. On some pages, there will be an
additional Math tab with mathematical details.

Column
-------------------------------------

### Marginal Effects

After running a regression model of some sort, the common way of interpreting the relationships between predictors and the outcome is by interpreting
the regression coefficients. In many situations these interpretations are straightforward, however, in some settings the interpretations can be
tricky, or the coefficients can simply not be interpreted.

These complications arise most commonly when a model involves an interaction or moderation. In these sort of models, interpretation of the
coefficients requires special care due to the relationship between the various interacted predictors, and the set of information obtainable from the
coefficients is often limited without resorting to significant algebra.

Marginal effects offer an improvement over simple coefficient interpretation. **Marginal means** allow you to estimate the average response under a
set of conditions; e.g. the average response for each racial group, or the average response when age and weight are at certain values.

**Marginal slopes** estimate the slope between a predictor and the outcome when other variables are at some specific value; e.g. the average slope on
age within each gender, or the average slope on age when weight is at certain values.

In either case, in addition to estimating these marginal effects, we can easily test hypotheses of equality between them via **pairwise
comparisons**. For example, is the average response different by ethnicity, or is the slope on age different between genders.

If you are familiar with interpreting regression coefficients, and specifically interactions, you may have some idea of how to start addressing all
the above examples. However, to complete answer these research questions would require more than simply a table of regression coefficients. With
marginal effects, one additional set of results can address all these questions.

Finally, marginal effect estimation is a step towards creating informative visualizations of relationships such as interaction plots.


# Data and Libraries

Sidebar {.sidebar}
-------------------------------------

```{r child = 'discussion_markdown/data.Rmd'}
```

Code {.tabset}
-------------------------------------

### Stata

```{r child = 'data.Rmd'}
```

### R Data

```{r child = 'R_markdown/data.Rmd'}
```

### R Packages

```{r child = 'R_markdown/packages.Rmd'}
```

# Categorical Variables {data-navmenu="Marginal Means"}

Sidebar {.sidebar}
-------------------------------------

```{r child = 'discussion_markdown/mm_cat.Rmd'}
```

Code {.tabset}
-------------------------------------

### Stata

```{r child = 'mm_cat.Rmd'}
```

### R (`emmeans`)

```{r child = 'R_markdown/mm_emmeans_cat.Rmd'}
```

### R (`ggeffects`)

```{r child = 'R_markdown/mm_ggeffects_cat.Rmd'}
```

### Math

```{r child = 'math_markdown/mm_cat_math.Rmd'}
```

# Interactions between Categorical Variables {data-navmenu="Marginal Means"}

Sidebar {.sidebar}
-------------------------------------

```{r child = 'discussion_markdown/mm_cat_ind.Rmd'}
```

Code {.tabset}
-------------------------------------

### Stata

```{r child = 'mm_cat_int.Rmd'}
```

### R (`emmeans`)

```{r child = 'R_markdown/mm_emmeans_cat_int.Rmd'}
```

### R (`ggeffects`)

```{r child = 'R_markdown/mm_ggeffects_cat_int.Rmd'}
```

### Math

```{r child = 'math_markdown/mm_cat_int_math.Rmd'}
```

# Continuous Variables {data-navmenu="Marginal Means"}

Sidebar {.sidebar}
-------------------------------------

```{r child = 'discussion_markdown/mm_con.Rmd'}
```

Code {.tabset}
-------------------------------------

### Stata

```{r child = 'mm_con.Rmd'}
```

### R (`emmeans`)

```{r child = 'R_markdown/mm_emmeans_con.Rmd'}
```

### R (`ggeffects`)

```{r child = 'R_markdown/mm_ggeffects_con.Rmd'}
```

### Math

```{r child = 'math_markdown/mm_con_math.Rmd'}
```

# Categorical and Continuous Interactions {data-navmenu="Marginal Means"}

Sidebar {.sidebar}
-------------------------------------

```{r child = 'discussion_markdown/mm_cat_con_int.Rmd'}
```

Code {.tabset}
-------------------------------------

### Stata

```{r child = 'mm_cat_con_int.Rmd'}
```

### R (`emmeans`)

```{r child = 'R_markdown/mm_emmeans_cat_con_int.Rmd'}
```

### R (`ggeffects`)

```{r child = 'R_markdown/mm_ggeffects_cat_con_int.Rmd'}
```

# Marginal Slopes {data-navmenu="Marginal Slopes"}

Sidebar {.sidebar}
-------------------------------------

```{r child = 'discussion_markdown/ms_con.Rmd'}
```

Code {.tabset}
-------------------------------------

### Stata

```{r child = 'ms_con.Rmd'}
```

### R (`emmeans`)

```{r child = 'R_markdown/ms_emmeans_con.Rmd'}
```

### Math

```{r child = 'math_markdown/ms_con_math.Rmd'}
```


# Categorical by continuous interaction {data-navmenu="Marginal Slopes"}

sidebar {.sidebar}
-------------------------------------

```{r child = 'discussion_markdown/ms_con_int.Rmd'}
```

Code {.tabset}
-------------------------------------

### Stata

```{r child = 'ms_con_int.Rmd'}
```

### R (`emmeans`)

```{r child = 'R_markdown/ms_emmeans_con_int.Rmd'}
```

### Math

```{r child = 'math_markdown/ms_con_int_math.Rmd'}
```

# Categorical Variables {data-navmenu="Pairwise Comparisons"}

Sidebar {.sidebar}
-------------------------------------

```{r child = 'discussion_markdown/pw_cat.Rmd'}
```

Code {.tabset}
-------------------------------------

### Stata

```{r child = 'pw_cat.Rmd'}
```

### R (`emmeans`)

```{r child = 'R_markdown/pw_emmeans_cat.Rmd'}
```

# Categorical Interactions {data-navmenu="Pairwise Comparisons"}

Sidebar {.sidebar}
-------------------------------------

```{r child = 'discussion_markdown/pw_cat_int.Rmd'}
```

Column {.tabset}
-------------------------------------

### Stata

```{r child = 'pw_cat_int.Rmd'}
```

### R (`emmeans`)

```{r child = 'R_markdown/pw_emmeans_cat_int.Rmd'}
```

# Marginal Slopes {data-navmenu="Pairwise Comparisons"}

Sidebar {.sidebar}
-------------------------------------

```{r child = 'discussion_markdown/pw_cat_con_int.Rmd'}
```

Column {.tabset}
-------------------------------------

### Stata

```{r child = 'pw_cat_con_int.Rmd'}
```

### R (`emmeans`)

```{r child = 'R_markdown/pw_emmeans_cat_con_int.Rmd'}
```

# Diff-in-diff {data-navmenu="Pairwise Comparisons"}

Sidebar {.sidebar}
-------------------------------------

```{r child = 'discussion_markdown/diff_in_diff.Rmd'}
```

Column {.tabset}
-------------------------------------

### Stata

```{r child = 'diff_in_diff.Rmd'}
```

### R (`emmeans`)

```{r child = 'R_markdown/diff_in_diff.Rmd'}
```

# Plotting

Sidebar {.sidebar}
-------------------------------------

```{r child = 'discussion_markdown/plotting.Rmd'}
```

Code {.tabset}
-------------------------------------

### Stata

```{r child = 'plotting.Rmd'}
```

### R (`interactions`)

```{r child = 'R_markdown/plotting_interactions.Rmd'}
```

# Odds Ratios {data-navmenu="Miscellaneous"}

Sidebar {.sidebar}
-------------------------------------

```{r child = 'discussion_markdown/oddsratios.Rmd'}
```

Code {.tabset}
-------------------------------------

### Stata

```{r child = 'oddsratios.Rmd'}
```

### R (`interactions`)

```{r child = 'R_markdown/oddsratios.Rmd'}
```
