There are two ways to do this. In the first, we can specify specific values of multiple variables, but only specify a single value per variable.

```{r}
summary(mod1 <- lm(y ~ sex + age + distance, data = m))
emmeans::emmeans(mod1, ~ 1, at = list(age = 30))
emmeans::emmeans(mod1, ~ 1, at = list(age = 40, distance = 10))
```

Note the use of `~ 1`. This argument (`spec`) cannot be excluded, but using `~ 1` basically ignores it and only focuses on the `at`.

Alternatively, we can specify it slightly differently, which allows for multiple values of a single variable, but not multiple variables.

```{r}
emmeans::emmeans(mod1, ~ age, cov.reduce = function(x) c(30, 40))
emmeans::emmeans(mod1, ~ age + distance, cov.reduce = function(x) c(30, 40))
```

Note that we're back to using `~ age`.

For either approach, we can add a categorical variable to the formula to examine across levels of that variable.

```{r}
emmeans::emmeans(mod1, ~ sex, at = list(age = 30))
emmeans::emmeans(mod1, ~ age + sex, cov.reduce = function(x) c(30, 40))
```