```{r}
library(haven)
library(ggeffects)
library(interactions)
```

The `ggeffects` package requires several other packages to work fully. However, it does not follow proper package framework and requires you to ensure you have the following packages installed as well:

- `effects`
- `emmeans`
- `ggplot2`

```{r}
library(effects)
library(emmeans)
library(ggplot2)
```

If you do not have those packages loaded, `ggeffects` will load them on demand (rather than at the time of loading `ggeffects`). If you do not have a required package installed, `ggeffects` will instead produce a warning, so keep an eye out for those and install packages as needed.

```{r}
m <- haven::read_dta("http://www.stata-press.com/data/r15/margex.dta")
m <- data.frame(m)
m$sex <- as_factor(m$sex)
m$group <- as_factor(m$group)
head(m)
```

A note on syntax: To make clear which function belongs to which library, I've used R's [double
colon](https://stat.ethz.ch/R-manual/R-devel/library/base/html/ns-dblcolon.html) notation to be explicit. You do not need to include
these. E.g. `library::function()` can be run simply as `function()`. There's a bit of possible confusion in that the workhorse function `ggeffect`,
belongs to the "ggeffects" package, so `ggeffects::ggeffect` can be run as just `ggeffect`.
