```{r}
summary(mod1 <- lm(y ~ class, data = m))
predictions(mod1, datagrid(class = unique))
marginalmeans(mod1, "class")
```

The `marginalmeans()` function is provided as a direct translation for `emmeans`
users. The `predictions()` function will be used more commonly so I recommend
becoming familiar with that.

The `datagrid()` function can also take in values. For example,

```{r}
predictions(mod1, datagrid(class = c(1, 2, 3)))
```

is an alternative to `class = unique` to specify the levels. (NOTE: These are
the "levels" not the values - it just happens that `class` has the same. See
below for what happens when it differs.) We can specify a subset of levels as
well.

```{r}
predictions(mod1, datagrid(class = c(1, 2)))
```

#### Factor labels

If your factor has labels which are not just the values, you need to refer to
those.

```{r, error = TRUE}
mod2 <- lm(y ~ sex, data = m)
str(m$sex)
predictions(mod2, datagrid(sex = c(1,2)))
predictions(mod2, datagrid(sex = c("male", "female")))
```
