There are several packages which can be used to estimate marginal means. So far,
this document implements `emmeans` and `ggeffects`, and I've begun writing up
`marginaleffects`.

[`emmeans`](https://cran.r-project.org/web/packages/emmeans/index.html) is based
on an older package, `lsmeans`, and is more flexible and powerful. However, it's
syntax is slightly strange, and there are often several ways to obtain similar
results which may or may not be the same. Additionally, while it's immediate
output is clear, accessing the raw data producing the output takes some doing.

[`ggeffects`](https://strengejacke.github.io/ggeffects/) is a package designed
to tie into the dplyr/ggplot universe. It's results are always a clean data
frame which can be manipulated or passed straight into ggplot. However,
`ggeffects` can *only* estimate marginal means, not marginal slopes or pairwise
comparisons. (If I'm wrong about that last statement, please let me know.)

[`marginaleffects`](https://vincentarelbundock.github.io/marginaleffects) is a
newer package which has a much more robust series of estimates. It focuses on
observation-level predictions which can then be averaged, rather than estimating
marginal effects directly. Theres often multiple ways to obtain the same result
which can be very frustrating. It's a very early version (
`r packageVersion("marginaleffects")` at time of last updating this document) so it
may see breaking changes.

The `marginaleffects` website has a [good write-up comparing between these
packages (as well as
Stata)](https://vincentarelbundock.github.io/marginaleffects/articles/alternative_software.html).
This is obviously written with bias towards being pro-`marginaleffects` so keep
that in mind while reading it.

In addition, the `interactions` package is used to plot interaction effects.

### `emmeans`

```{r}
library(emmeans)
```

### `ggeffects`

The `ggeffects` package requires several other packages to work fully. However,
it does not follow proper package framework and install them for you, rather it
requires you to ensure you have the following packages installed as well:

- `effects`
- `emmeans`
- `ggplot2`

```{r}
library(ggeffects)
library(effects)
library(emmeans)
library(ggplot2)
```

If you do not have those packages loaded, `ggeffects` will load them on demand
(rather than at the time of loading `ggeffects`). If you do not have a required
package installed, `ggeffects` will instead produce a warning, so keep an eye
out for those and install packages as needed.

### `marginaleffects`

```{r}
library(marginaleffects)
```

(Using `marginaleffects` remains a work-in-progress.)

### `interactions`

```{r}
library(interactions)
```

### Software Versions

This document was last compiled using `r system("egrep 'Stata.*[0-9.]{4}' /Applications/Stata/StataSE.app/Contents/Info.plist | awk -F '>' '{print $2}' | awk -F '<' '{print $1}'", intern = TRUE)` and R `r getRversion()`.

R package versions at last compilation:

| Package           | Version                               |
|-------------------|---------------------------------------|
| `emmeans`         | `r packageVersion("emmeans")`         |
| `ggeffects`       | `r packageVersion("ggeffects")`       |
| `effects`         | `r packageVersion("effects")`         |
| `marginaleffects` | `r packageVersion("marginaleffects")` |
| `ggplot2`         | `r packageVersion("ggplot2")`         |
| `interactions`    | `r packageVersion("interactions")`    |
